Creative Destruction

### 2016-03-18 paper.tex
asdf

The supply and demand theory has had strong explanatory power over the years, but it doesn't expalin one crucial observation: economies grow.
In a neoclassical economic framework, an economy reaches equilibrium and stabilizes.
Once an economy stabilizes, the only for it to grow is via exogenous inputs, for example, imperialism, increased supply of raw materials.

@@

This the idea of creative destruction.
The old displaces new; there is not room for both of them.
Marx and Schumpeter had interesting concern: what if there is too much destruction by new technology such that society destroys itself.
The capitalist system is a cycle of creation and destruction, and it seems possible, on a theoretical level, that the destruction may overwhelm the construction, spelling disaster for society.
While on a social level this is hard to imagine, it is somewhat easier to fathom on technological level.

This paper will explore the idea of creative destruction and its philosophical implications on technology. 
To whet the appetite, a few questions that will be addressed are:
\begin{itemize}
    \item New technology technically does not ``destroy'' old technology; rather, it makes the old technology lose economic value.
    \item But, what effect does new technology have on old technology? Does it make it obsolete? Does the ``idea'' of the old technology change? Does the ``meaning'' of technology, which is contextual and socially constructed, changed due to a change in society or due to a change in technology? That question implies that technosphere, as one might call it, is distinct from society, which social constructs persuasively argue is not necessarily the case.
    \item Is technology still cumulative? Or in what nuanced sense is it cumulative?
    \item How does this idea fit into evolution?
    \item What role does this play in selection?
    \item Any effect on combination? gut says no, but not sure
\end{itemize}

@@

Beyond innovation, they are interested in knowledge spillovers, believing that increases in public knowledge contribute to economy wide returns.
They distinguish between knowledge and technological obsolescence and creative destruction:
\begin{displayquote}
``Old knowledge eventually is made obsolete by the emergence of newer, superior knowledge. We call this phenomenon `knowledge' or `technological' obsolescence, and distinguish it from the obsolescence in value represented by creative destruction.
That is, new ideas have two distinct effects on the current stock of ideas. 
They make the products represented by those ideas less valuable (creative destruction or value obsolescence), and they make the knowledge represented by those ideas less relevant in the production of new knowledge (knowledge or technological obsolescence). 
The strength of knowledge spillovers, and hence the growth of the economy, will depend on the parameters of the processes of knowledge diffusion and knowledge obsolescence.'' (92)
\end{displayquote}
Creative destruction is specifically the effect of new innovation on the economic ``value'' of products.

From a philosophy perspective, it seems overly limiting to only consider creative destruction as the impact that new creations have on the economic value of old technology.
As a contrived example, consider Intel who sells CPUs. 
If Intel creates

@@

First, creative destruction makes technology more unpredictable.
The reason for this hearkens back to Brian Arthur's idea of primitives. 
Arthur reasons that new technology is the product of a new combination of primitives. 
Creative destruction

>>>>

The supply and demand theory doesn't explain one crucial observation: economies grow.
In a neoclassical economic framework, an economy reaches equilibrium and stabilizes.
Once an economy stabilizes, the only for it to grow is via exogenous inputs, for example, imperialism, increased supply of raw materials.

@@

This the idea of creative destruction.
The old displaces new; the better, more efficient verison makes the old, costly version unviable.
For example, the smartphone, sophisticated little computer with the power the text and call, destroyed the economic viability of the dumb-phone.
Marx and Schumpeter were concerned about the possible effects of creative destruction: what if there is too much destruction?
The capitalist system is a cycle of creation and destruction, and it seems possible, on a theoretical level, that the destruction may overwhelm the construction, spelling disaster for society.

@@

Beyond innovation, they are interested in knowledge spillovers, believing that increases in public knowledge contribute to economy wide returns.
They distinguish between knowledge and technological obsolescence and creative destruction:
\begin{displayquote}
``Old knowledge eventually is made obsolete by the emergence of newer, superior knowledge. We call this phenomenon `knowledge' or `technological' obsolescence, and distinguish it from the obsolescence in value represented by creative destruction.
That is, new ideas have two distinct effects on the current stock of ideas. 
They make the products represented by those ideas less valuable (creative destruction or value obsolescence), and they make the knowledge represented by those ideas less relevant in the production of new knowledge (knowledge or technological obsolescence). 
The strength of knowledge spillovers, and hence the growth of the economy, will depend on the parameters of the processes of knowledge diffusion and knowledge obsolescence.'' (92)
\end{displayquote}
Creative destruction is specifically the effect of new innovation on the economic ``value'' of products.
Caballero and Jaffe measures the effect on innovation on industries by using market value and patent data on 567 large U.S. firms.
They found that:
\begin{displayquote}
    ``In an average sector at an average year a firm that does not invent sees it relative value to that of the industry erode by about 4\%.''
\end{displayquote}
That is, there is approximately 4\% destruction of pre-existing technology on average per year. 
Different industries observed varying levels of destruction.
On the high end, the pharmaceutical industry observed 25\% destruction per year, yet multiple sectors observed 0\% destruction.

We find Caballero and Jaffe's method to be interesting and the results to be useful, but the story is not yet complete.
The model that Caballero and Jaffe use to compute creative destruction\footnote{The model is not communicated in this paper} is complicated and computed over a short number of years.
However, the findings are consistent with intuition.
4\% destruction on average per year is a reasonable number, and it is to be expected that pharmaceuticals see a high level destruction relative to other industries. 
Further research could be done on the economic side of creative destruction.
First, Jaffe and Caballero could further investigate the nature of the technologies that are being destroyed, charactering the nature of the destruction (e.g. is it intra-firm created, extra-firm created, expected destruction, unexpected destruction, and so on.).
Second, it would be interesting to measure creative destruction using other methods, for example using research and development expenditures.
Third, it would be interesting to see how the industries are effected by the varying degrees of creative destruction, and the way that the firms innovate.

We greatly appreciate Caballero and Jaffe's results, and use them to motivate our research and argue about the effects of creative destruction.

@@

First, creative destruction makes technology more unpredictable.
The reason for this hearkens back to Brian Arthur's idea of primitives. 
Arthur reasons that new technology is the product of a new combination of tecnological primitives. 
At a given time, there exists a set of primitives. 
A new technology is created out of recombining the existing primitives, that the new technology is added to set of the primitives.
Creative destruction suggests that when a new technology is created and added to the set of primitives, pre-existing technologies may be removed from the set of primitives.
The primitives interact with each other - some in a complementary way and some in a supplementary way (like creative destruction). 
When one is trying to predict the future path of technology, the primitives available in the future may be missing primitves that are present now, making the task substantially more complex.

Second, creative destruction does introduce a feeling that technology is dangerous.
Mark and Schumpeter worried that society would create so much that it would consume and destroy itself, and this is a possibility with technology.
In section 2, we outlined an example of how the internet may be destroyed if there is too much innovation too fast.
New technology, introduced and implmented in the wrong way, could undermine the technological infrastructure that has been built up. 
With the technological infrastructure gone, it's not difficult to imagine the negative social reprecussions.

Third, we have the perhaps unfortunate situation where there is constantly new technology.
New iphones, new computers, new drugs.
With the creation of each new invention, creative destruction theory says that old technology may be destroyed.
The old technology, which was likely less expensive (for various reason), is now inaccessible: all that exists is the new, expensive technology.
The implication is that the poor, who cannot afford the new technology, are left with nothing. 
In cases, the poor can't afford the new technology, and the old technology, which they may or may not be able to afford, no longer exists.

The most pressing area where we see this problem is the pharmaceutical industry.
There are two phases for a drug on the market. 
The first phase is where the drug is patented, the firm has a monopoloy on it and prices are high.
The second phase is after the patent expires, and the drug enters the public domain. 
Unfortunatley, some drugs don't benefit from the second phase. 
\footnote{I did not look up examples and examine this in detail.}
There are cases where people don't benefit from drugs being the second phase.
It is not economically viable for firms to mass produce the drug, so poorer people can't access it.

\section{Conclusion}
Creative Destruction, the idea that new technologies make existing technologies less relevant, fits in nicely with the technological theory of evolution.
In particular, creative destruction explains a portion of the selection process, and how new technologies affect the fitness landscape of existing technologies.
Moreover, creative destruction has a strong analog in biological evolution, competitive exclusion.
Creative destruction and competitive exclusion share many commonalities and few differences, but the similarities suggest that at least some part of technological selection is quite similar to biological selection.

Creative destruction also brings three philosophical considerations to the table: (1) technology is complex and unpredictable, (2) technology is potential dangerous to itself and (3) the nature of technology is classist, harming the poor in some instances.
We consider the third point to be most important and most easily observed.
Technology has the power to do great good and great harm. 
The good and harm can stem from individual technologies themselves, like a technology that cures cancer is great, but drones that kill humans are arguably bad. 
Zooming out, the nature of technology and how it evolves has a devestating effect on the poor.
The poor cannot afford the new technologies, and the new technologies often cause firms to not produce old technologies because they are not economically viable, leaving the poor without anything. 
Here, it is the process by which technology is created and evolves which harms the poor, not the technology itself.
